---
title: "Project_2: Data Transformation"
author: "Alyssa Gurkas, Alina Vikhevich, Musrat Jahan, Olivia Azevedo"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load libraries}
# load libraries
library(RSocrata)
library(tidyverse)
library(readr)
library(readxl)
library(janitor)
library(lubridate)
```

```{r call-dart-api}
# API call to NYS DEC DART
# would be nice to call the API for NPDES permits (as listed in p_type)
# dart <- read.socrata("https://data.ny.gov/resource/mbk7-f2r2.json")
# dart <- read_csv("C:/Users/AGURKAS/OneDrive - Environmental Protection Agency (EPA)/Profile/Downloads/dart_3.7.2025.csv")
# dart <- dart |> 
# mutate(date_received=as.Date(date_received, format = "%d-%m-%Y")) |> 
# filter(date_received > '2019-12-31')
# write_csv(dart,"dart_2020_2025.csv")

read_csv("dart_2020_2025.csv")

```

```{r defining-regex-patterns}
# p_type = permit type, in this exercise we are filtering for wastewater permits
p_type <- c("P/C/I SPDES - Surface Discharge",
            "Municipal SPDES - Surface Discharge",
            "Industrial SPDES - Surface Discharge")

# defining the regex patterns for the IDs we want to track
npdes_pattern <- "NY\\d{7}"
gp_pattern <- "GP\\d{7}"
individual_pattern <- "NY[A-Z]\\d{2}[A-Z]\\d{3}"
all_patterns <- paste(npdes_pattern,gp_pattern,individual_pattern, sep="|")
```

```{r creating-universe-v1}
# creating the permit universe pulling from NYSDEC's DART System 
# and detecting the string patterns within DART to assign permit type: npdes, 
# individual(i.e., a permit covered under a general permit), general, 
# or multi (meaning the DART entry had multiple associated IDs).

universe <- dart |>  
  filter(`permit_type` %in% p_type) |>  
  mutate(
    npdes = str_count(`other_known_ids`, npdes_pattern),
    individual = str_count(`other_known_ids`, individual_pattern),
    gp =  str_count(`other_known_ids`, gp_pattern),
    sum_ids = rowSums(across(c(`npdes`, `individual`,`gp`))),
    npdes_id = str_extract_all(`other_known_ids`, all_patterns),
    date_received=as.Date(date_received, format = "%d-%m-%Y")
    ) |> 
  mutate(applicant_id =cur_group_id(),.by = applicant) |> # creating applicant id
  mutate(facility_id = cur_group_id(),.by = c(facility,location,town_or_city)) |> # creating facility id
  distinct() |> # removing duplicate rows 
  mutate(
          dart_p_type = case_when(sum_ids  > 1 ~ "multi",
                               sum_ids & npdes == 1 ~ "npdes",
                               sum_ids & individual == 1 ~ "individual",
                               sum_ids & gp == 1 ~ "gp")) |>  
  unnest_longer(npdes_id, keep_empty = FALSE) |> 
  filter(!is.na(npdes_id))

```
Note: The code above filters entries that did not have a NPDES ID listed in the 
"Other Known IDs" column, however, were listed as NPDES permits in the Permit 
Type Column. However, out of 35,642 entries, only 69 were missing NPDES IDs. 

```{r note-on-str-count}
# the str_count statements are taking a sum of associated permits from DART data 
# entries. If a permit is associated with multiple ids it is assigned multi, 
# if it is a regular npdes permit, it is called npdes. If the permit is an 
# individual permit under a general permit, it is called individual. if it is a 
# general permit, it is called 'gp'. for information about permit types please 
# see the "What is the process for applying for coverage under a NPDES general 
# permit" section of the EPA NPDES Permit Basics Page.
# https://www.epa.gov/npdes/npdes-permit-basics
```



```{r tbl1-permit-lvl}
# this table shows the most recent permit information
tbl1_permit_lvl <- universe  |> 
  group_by(npdes_id) |> 
  slice(which.max(date_received)) |>
  select(npdes_id,facility_id,application_id,applicant,applicant_id,permit_type,
         status,date_received,upa_class,seqr_class,seqr_determination,
         lead_agency,coastal_zone_status, final_disposition,permit_effective_date,
         permit_expration_date,dec_contact,shpa_status,enivronmental_justice)
```

```{r tbl2-permit-action-lvl}
# this table shows the permit history. each observation in this table represents
# a permit action.

# if there are true duplicates (meaning all rows are equal, then the duplicate
# row can be dropped)

tbl2_permit_act_lvl <- universe |> 
    mutate(action_id = paste(npdes_id,date_received, sep = "_")) |> 
    distinct() |> 
    mutate(dup_flag = duplicated(action_id),
           transfer_flag=str_detect(toupper(short_description),"TRANSFER")) |> 
  select(action_id,facility,facility_id,npdes_id,application_id,applicant,
         application_type,date_received,status,short_description,
         enb_publication_date,written_comments_due,dup_flag,transfer_flag)
```

```{r tbl3-facility-table}
# this table shows the facility information. Each observation in this table
# represents a facility associated with NPDES permits.
tbl3_facility_lvl <- universe |> 
  select(facility_id,facility,
         location,town_or_city) |> 
  distinct() |> 
  arrange(facility_id)
```

```{r tbl4-applicant-table}
# this table shows the applicant information. Each observation in this table
# represents a permit applicant for NPDES permits.
tbl4_app_lvl <- universe |> 
  group_by(applicant_id) |> 
  slice(which.max(date_received)) |>
  select(applicant_id,applicant,application_id)
```

Questions to answer:
- which facilities have multiple NPDES permits associated with it?
- which applicants are associated with multiple NPDES permits?
- # how do we identify if a facility has numerous permits?

        (1) tbl1 - permit table: the purpose of this table is to have the most
        recent permit information. This will have one row per permit.
        
        (2) tbl2 - permit action table: the purpose of this table is to have a 
        table with every permit-action. This means there should be one row per
        permit action. 
        
        (3) tbl3 - facility table: the purpose of this table is to have 
        information on the facility. 
        
        (4) tbl4 - applicant table: the purpose of this table is to have 
        information about the applicant.  

Considerations:
- There was missing data, such as NPDES IDs. This means that some permit 
information may not be available. 

- There may be facilities that are listed as different facilities due to address
changes. This information should be paired with other facility information, like
EPA's Enforcement and Compliance History Online 
(ECHO, available at: https://echo.epa.gov/).

- For entries that were made on the same day for a particular permit, it is not 
possible to identify which entry was made first. Permit transfer actions are 
largely affected by this. Due to this, duplicates and transfers are flagged for 
manual review. 



