---
title: "Insert Here"
author: "Alyssa Gurkas"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
    number_sections: false
    toc_collapsed: true
    toc_depth: 4
---

[Github Repository](https://github.com/amgurkas/data-acquisition/tree/main/Week_9/) 

```{r set-chunk-options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,      
  results = "hide", 
  warning = FALSE,  
  error = FALSE,   
  message = FALSE)
```

```{r load-libraries}
library(tidyverse)
library(nytimes)
library(httr)
library(jsonlite)
```

## "Manual" API Call using the `httr` and `jsonlite` packages. 
```{r manual-api-call}
# Retrieve API key from environment
api_key <- Sys.getenv("NYTIMES_API_KEY")

# Save the Top Sports Stories URL with the API Key included
url <- paste0("https://api.nytimes.com/svc/topstories/v2/sports.json?api-key=", 
              api_key)

# Fetch data
response <- GET(url)
# Check response type
http_type(response)
# Check for errors
http_error(response)
# Convert data from JSON to Text
data <- fromJSON(content(response, as="text"))

# Create sample dataframe of results
articles <- data$results %>%
  select(title, abstract, url, published_date)
```

## [NYTimes Package](https://github.com/news-r/nytimes)
Another way to retrieve the data from the NYTimes API call is using their 
R Package, and the function ny_archive(). 
```{r nytimes-package}
# get all articles from last five years
archive <- ny_archive(2024, 7)

# displaying the structure of the archive obj
str(archive[1:5]) 
```

```{r keyword-ele-analysis}
# displaying the structure of the first keywords element within the obj archive
str(archive[[1]]$keywords)

# Check if ESPN is a keyword in the first article using a grepl statement
any(grepl("ESPN", archive[[1]]$keywords))

# Test if ESPN is a keyword in the first article using stringr function
any(str_detect(unlist(archive[[1]]$keywords),"ESPN"))

# Create a test function that returns TRUE if 'ESPN' is in the keywords
contains_ESPN <- function(article) {
  any(str_detect(str_to_upper(unlist(archive[[article]]$keywords)),"ESPN"))
}

# Test the function with the first article in the archive
contains_ESPN(archive[[1]])  # Should return TRUE or FALSE
```

```{r keyword-funct}
# create a function that will search through the archive list
# and subset articles containing the keyword "ESPN"
filter_espn_articles <- function(article_index) {
  sublist <- archive[[article_index]]
  news_desk <- sublist$news_desk
  keywords_list <- sublist[["keywords"]]
  contains_espn <- any(grepl("espn", keywords_list, ignore.case = TRUE))
  return_df <- data.frame(article_index=article_index,
                          news_desk=news_desk,
                          contains_espn=contains_espn)
  return(return_df)
}

result <- map_df(
                seq_along(archive),
                filter_espn_articles
                )
```

```{r sports_nd_df}
# sports_nd <- result |> 
#   filter(news_desk=="Sports")
```


**Potential Analysis for ESPN-Related Insights**
*Frequency Analysis*: How often does ESPN appear in NYT sports articles?
*Sentiment Analysis*: What is the sentiment of ESPN-related coverage?
*Topic Modeling*: Identify trending topics associated with ESPN.


