---
title: "Assignment 3"
author: "Alyssa Gurkas"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

# Understanding Database Normalization
Database normalization is a method to organize data in a consistent way. 
Normalizing data can help ensure there are not duplicate entries and maintain 
data integrity. Using this method, the data is organized in many tables, linked
together by relationships (such as a unique id or key). There are a few levels
to normalized datasets: 1NF, 2NF, 3NF. 

```{r load-librarys, include=FALSE,message=FALSE, warning=FALSE, error=FALSE}
library(tidyverse)
library(RSocrata)
library(janitor)
library(knitr)
library(kableExtra)
```

<br> 
If a database were to be considered 1NF, or in the first normal form, it would 
be meet the following criteria:

```{=html}
<style>
.blue-text { color: #0077B6; }
</style>
```
1.  [Each cell does not hold more than one value]{.blue-text}\
2.  [Every row has a unique identifier]{.blue-text}\
3.  [Each column only has one value for each row in the table]{.blue-text}

<br>

An example dataset that is in the 1NF form is New York State's Energy Efficiency
Completed Projects Dataset:
```{r data-connection, include=TRUE,message=FALSE, warning=FALSE, error=FALSE}
# completed energy efficiency projects in NY 
# documentation here: https://data.ny.gov/Energy-Environment/Energy-Efficiency-Completed-Projects-Beginning-198/erjw-j2zx/about_data

# loading data from open ny data
ny_ee_proj <- read.socrata("https://data.ny.gov/resource/erjw-j2zx.json") 

# calculating the number of unique project codes
n_distinct(ny_ee_proj$project_code)

# calculating the number of rows
nrow(ny_ee_proj)

# creating a data frame with an agency ID, and rearranging the columns
df_1nf <- ny_ee_proj %>% 
  group_by(agency) %>% 
  mutate(agency_id =cur_group_id()) %>% 
  ungroup() |> 
  relocate(`project_code`,`project_name`,`year_completed`,`date_completed`,`agency`,`agency_id`)

df_1nf_clean <- clean_names(df_1nf, "title") |> 
  rename("County" = "Ny State County")

# rendering the 1NF data frame, showing the first 6 rows and columns
knitr::kable(head(df_1nf_clean, format ="markdown", n = c(6, 6))) 
```

The data seen above is managed by The New York Power Authority, and is
collected from project contracts, including project name and codes, location,
cost estimates, and energy efficiency savings estimates. This collection effort
is related to the Energy Efficiency Market Acceleration Program (EE MAP), which
was an initiative lead by former Governor Cuomo to advance market development
and commercialization of new energy efficient technologies.

<br>
To transform the data into the 2NF form, the code below creates two different 
tables, the Agency Table, and the Energy Project Table:
```{r 2NF, include=TRUE, message=FALSE, warning=FALSE, error=FALSE}
# agency table
# removing values where Agency title is empty (NA)
df_2nf_ex1 <- df_1nf %>% 
  select(agency_id,agency,address,city,zip,address_ext) %>% 
  distinct() %>% 
  drop_na(agency) %>% 
  arrange(agency_id) |> 
  clean_names("title")

# Energy Project Table
df_2nf_ex2 <- df_1nf %>% 
  select(-agency,-address,-city,-zip,-address_ext) %>% 
  relocate(agency_id) |> 
  clean_names("title")

# rendering the Agency Table, showing the first 6 rows and columns
knitr::kable(head(df_2nf_ex1, format ="markdown", n = c(6, 6)))

# rendering the Energy Project Table, showing the first 6 rows and columns
knitr::kable(head(df_2nf_ex2, format ="markdown", n = c(6, 6)))
```
<br>

By adding an Agency ID, it ensures that if there were two organizations with 
the same name, or similar names, there would still be a distinct key to identify
them. Additionally, with a unique ID it can become easier to look up information
about an agency:

```{r agency-look-up, message=FALSE, warning=FALSE, error=FALSE}
df_2nf_ex2 |> 
  filter(`Agency Id` == 1) |> 
  head(n = c(6, 6))
```

Below is an example of the data as unnormalized: 
```{r unnormalized-ex1, include=TRUE, message=FALSE, warning=FALSE, error=FALSE}

# creating a function to combine the year_completed and date_completed columns 
# as a list in a new column titled 'Completion'
df_1nf$Completion <- mapply(function(year, date) list(Year = as.integer(year), 
                                                      Date = as.Date(date, 
                                                                     format = "%Y-%m-%d")),
                            df_1nf$year_completed, 
                            as.Date(df_1nf$date_completed, format = "%Y-%m-%d"),
                            SIMPLIFY = FALSE)

df_un <- df_1nf[, c("project_code", "Completion")]

# render the unnormalized dataframe
kable(head(df_un, format = "html", escape = FALSE)) |> 
 kable_styling("striped", full_width = FALSE)

```

In this example, the completion dates and year is combined into one column. As a
note, when the table is rendered, it calculates the number of days since 
January 1, 1970 (representing when R was created). This type of data structure
could complicate data entry or analysis, as the dates may need to be extracted
from the `Completion` column to be analyzed, and if a date needed to be edited,
it would not be simple to do so. 

<br> 

To achieve an 3NF form, the database should be in the second form (2NF), and
have no transitive partial dependency. This means there are no
duplicates or transitory keys. For example, the code below transforms the data 
to the third form:

```{r 3NF, include=TRUE,message=FALSE, warning=FALSE, error=FALSE}
# creating project code table
df_3nf_ex1_projectcode <- df_2nf_ex2 |>
  select(`Project Code`, `Project Name`) |>
  distinct() |> 
  clean_names("title")

# creating the project table
df_3nf_ex2_projects <- df_2nf_ex2 |>
  select(-`Project Name`) |> 
  clean_names("title")

# render the project code table
knitr::kable(head(df_3nf_ex1_projectcode, format ="markdown"))

# render the project table with descriptive columns
knitr::kable(head(df_3nf_ex2_projects, format ="markdown", n = c(6, 6)))
```

# Character Manipulation 
In this section of the assignment, using 538's data from the article, [The Economic Guide To Picking A College Major](https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/), we will identify which majors contain the terms "data" or "statistics". 

<br>

To do this, first, you should read in the data, which is hosted on [538's Github Page](https://github.com/fivethirtyeight/data/blob/master/college-majors/majors-list.csv).
```{r load-college-majors, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
majors_data <- read_csv("https://raw.githubusercontent.com/fivethirtyeight/data/refs/heads/master/college-majors/majors-list.csv")
```

Then, using the stringr package, we can filter the data to only include majors that incorporate the terms "data" or "statistics": 
```{r str-detect-ex, include=TRUE, message=FALSE, warning=FALSE, error=FALSE}
majors_data_filtered <- majors_data %>% 
  filter(str_detect(Major,"DATA|STATISTICS"))
```

This creates a dataframe with three majors:
```{r data-stat-majors, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
majors_data_filtered$Major
```

## Regex Part 1: 
In the assignment instructions, there is an exercise to translate the following
regular expressions in layman's terms:

1. (.)\1\1

2.  "(.)(.)\\2\\1"

3. (..)\1

4. "(.).\\1.\\1"

5.  "(.)(.)(.).*\\3\\2\\1"

<br>

## Answers: 

1. The "(.)" signifies a group, "\1" signifies a character, and the second "\1"
signifies a matching character. This means that a pattern following this regular
expression could be "aaa", "111" or "...".
2. This regular expression means there are two groups, depicted by the "(.)"
symbols. The "\2" means the exact same character that appeared in the
second group appears again. So, an example expression that matches this pattern
could be "ceec", "1221", or "woow".
3. The (..) means there is a group with two characters. The "\1" means that it 
matches two characters. Therefore, this pattern matches two consecutive 
identical pairs of characters. This could be "abab", "3434", or "aaaa". 
4. "(.).\\1.\\1" is a pattern where there is a repeating character that appears
three times with a character in between each repetition. A string following this
pattern is "1_1_1", or "abaca".
5. This pattern can be explained as one where there are three groups, that occur
twice, however the second time the group appears, it is mirrored. An example of 
a string that matches this pattern is "123xxxx321".

## Regex Part 2
In this part of the assignment, there are instructions to construct regular 
expressions to match words that:

1. Start and end with the same character.
2. Contain a repeated pair of letters (e.g. "church" contains "ch" repeated 
twice.)
3. Contain one letter repeated in at least three places (e.g. "eleven" contains 
three "e"s.)

## Answers
1. ^(.).*\\1$
2. \b(..).*?\1\b
3. \b.*([a-zA-Z]).*\1.*\1.*\b

