---
title: "Assignment_3"
author: "Alyssa Gurkas"
date: "`r Sys.Date()`"
output: html_document
---

# Understanding Database Normalization
Database normalization is a method to organize data in a consistent way. 
Normalizing data can help ensure there are not duplicate entries and maintain 
data integrity. Using this method, the data is organized in many tables, linked
together by relationships (such as a unique id or key). There are a few levels
to normalized datasets: 1NF, 2NF, 3NF. 

```{r load-librarys, include=FALSE}
library(tidyverse)
library(RSocrata)
library(DT)
library(janitor)
```

<br> 
If a database were to be considered 1NF, or in the first normal form, it would 
be meet the following criteria:
1. Each cell does not hold more than one value
2. Every row has a unique identifier
3. There are no duplicated entries
4. Each column only has one value for each row in the table
<br>

An example dataset that is in the 1NF form is New York State's Energy Efficiency
Completed Projects Dataset:
```{r data-connection, include=FALSE}
# completed energy efficiency projects in NY 
# documentation here: https://data.ny.gov/Energy-Environment/Energy-Efficiency-Completed-Projects-Beginning-198/erjw-j2zx/about_data
ny_ee_proj <- read.socrata("https://data.ny.gov/resource/erjw-j2zx.json") 

n_distinct(ny_ee_proj$project_code)
nrow(ny_ee_proj)

df_1nf <- ny_ee_proj %>% 
  group_by(agency) %>% 
  mutate(agency_id =cur_group_id()) %>% 
  ungroup()

df_1nf_clean <- clean_names(df_1nf, "title")
```

```{r dt-1NF, echo=TRUE}
datatable(head(df_1nf_clean,
              class = c('cell-border stripe','compact')))
```

<br>
An example of a dataset that is in the 2NF form is the same dataset, 
but transformed:
```{r 2NF, include=TRUE}
# agency table
# removing values where Agency title is empty (NA)
df_2nf_ex1 <- df_1nf %>% 
  select(agency_id,agency,address,city,zip,address_ext) %>% 
  distinct() %>% 
  drop_na(agency) %>% 
  arrange(agency_id) 

# Energy Project Table
df_2nf_ex2 <- df_1nf %>% 
  select(-agency,-address,-city,-zip,-address_ext) %>% 
  relocate(agency_id)
```

