---
title: "Assignment_3"
author: "Alyssa Gurkas"
date: "`r Sys.Date()`"
output: html_document
---

# Understanding Database Normalization
Database normalization is a method to organize data in a consistent way. 
Normalizing data can help ensure there are not duplicate entries and maintain 
data integrity. Using this method, the data is organized in many tables, linked
together by relationships (such as a unique id or key). There are a few levels
to normalized datasets: 1NF, 2NF, 3NF. 

```{r load-librarys, include=FALSE}
library(tidyverse)
library(RSocrata)
library(DT)
library(janitor)
```

<br> 
If a database were to be considered 1NF, or in the first normal form, it would 
be meet the following criteria:
1. Each cell does not hold more than one value
2. Every row has a unique identifier
3. There are no duplicated entries
4. Each column only has one value for each row in the table
<br>

An example dataset that is in the 1NF form is New York State's Energy Efficiency
Completed Projects Dataset:
```{r data-connection, include=FALSE}
# completed energy efficiency projects in NY 
# documentation here: https://data.ny.gov/Energy-Environment/Energy-Efficiency-Completed-Projects-Beginning-198/erjw-j2zx/about_data
ny_ee_proj <- read.socrata("https://data.ny.gov/resource/erjw-j2zx.json") 

n_distinct(ny_ee_proj$project_code)
nrow(ny_ee_proj)

df_1nf <- ny_ee_proj %>% 
  group_by(agency) %>% 
  mutate(agency_id =cur_group_id()) %>% 
  ungroup()

df_1nf_clean <- clean_names(df_1nf, "title")
```

```{r dt-1NF, echo=TRUE}
datatable(head(df_1nf_clean,
              class = c('cell-border stripe','compact')))
```

<br>
An example of a datasets that are within the 2NF form is the same dataset, 
but transformed:
```{r 2NF, include=TRUE}
# agency table
# removing values where Agency title is empty (NA)
df_2nf_ex1 <- df_1nf %>% 
  select(agency_id,agency,address,city,zip,address_ext) %>% 
  distinct() %>% 
  drop_na(agency) %>% 
  arrange(agency_id) 

# Energy Project Table
df_2nf_ex2 <- df_1nf %>% 
  select(-agency,-address,-city,-zip,-address_ext) %>% 
  relocate(agency_id)
```


<br> 

To achieve an 3NF form, the database should be in the second form (2NF), and
have no transitive partial dependency. This means [insert here]. For example,
the code below transforms the data to the third form:
```{r 3NF, include=TRUE}
# here will have code that transforms the data to the 3rd form

```

# Character Manipulation 
In this section of the assignment, using 538's data from the article, [The Economic Guide To Picking A College Major](https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/), we will identify which majors contain the terms "data" or "statistics". 

<br>

To do this, first, you should read in the data, which is hosted on [538's Github Page](https://github.com/fivethirtyeight/data/blob/master/college-majors/majors-list.csv).
```{r load-college-majors, include=TRUE, messages=FALSE}
majors_data <- read_csv("https://raw.githubusercontent.com/fivethirtyeight/data/refs/heads/master/college-majors/majors-list.csv")
```

Then, using the stringr package, we can filter the data to only include majors that incorporate the terms "data" or "statistics": 
```{r str-detect-ex, include=TRUE}
majors_data_filtered <- majors_data %>% 
  filter(str_detect(Major,"DATA|STATISTICS"))
```

This creates a dataframe with three majors:
```{r data-stat-majors, include=TRUE, echo=TRUE}
majors_data_filtered$Major
```

## Regex Part 1: 
In the assignment instructions, there is an exercise to translate the following
regular expressions in layman's terms:
1. (.)\1\1
2.  "(.)(.)\\2\\1"
3. (..)\1
4. "(.).\\1.\\1"
5.  "(.)(.)(.).*\\3\\2\\1"

<br>

## Answers: 

1. The "(.)" signifies a group, "\1" signifies a character, and the second "\1"
signifies a matching character. This means that a pattern following this regular
expression could be "aaa", "111" or "...".
2. This regular expression means there are two groups, depicted by the "(.)"
symbols. The "\2" means the exact same character that appeared in the
second group appears again. So, an example expression that matches this pattern
could be "ceec", "1221", or "woow".
3. The (..) means there is a group with two characters. The "\1" means that it 
matches two characters. Therefore, this pattern matches two consecutive 
identical pairs of characters. This could be "abab", "3434", or "aaaa". 
4. "(.).\\1.\\1" is a pattern where there is a repeating character that appears
three times with a character in between each repetition. A string following this
pattern is "1_1_1", or "abaca".
5. This pattern can be explained as one where there are three groups, that occur
twice, however the second time the group appears, it is mirrored. An example of 
a string that matches this pattern is "123xxxx321".

## Regex Part 2
In this part of the assignment, there are instructions to construct regular 
expressions to match words that:

1. Start and end with the same character.
2. Contain a repeated pair of letters (e.g. "church" contains "ch" repeated 
twice.)
3. Contain one letter repeated in at least three places (e.g. "eleven" contains 
three "e"s.)

## Answers
1. ^(.).*\\1$
2. \b(..).*?\1\b
3. \b.*([a-zA-Z]).*\1.*\1.*\b

