---
title: "Sentiment Analysis"
author: "Alyssa Gurkas"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
    number_sections: false
    toc_collapsed: true
    toc_depth: 4
    code_folding: show
---

[Github Repository](https://github.com/amgurkas/data-acquisition/tree/main/Week_10/) 
<br>
Source: TweetNLP: Cutting-Edge Natural Language Processing for Social Media”. Jose Camacho-Collados, Kiamehr Rezaee, Talayeh Riahi, Asahi Ushio, et al. Proceedings of the 2022 Conference on Empirical Methods in Natural Language Processing (EMNLP): System Demonstrations [Link paper](https://arxiv.org/abs/2206.14774)

```{r set-chunk-options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,      
  results = "hide", 
  warning = FALSE,  
  error = FALSE,   
  message = FALSE)
```

### Load Libraries
```{r load-library}
library(tidyverse)
library(httr)
library(jsonlite)
library(sentimentr)
library(janitor)
```

### Connecting to Twitter API 
The code below connects to the Twitter API and returns posts from the last 
seven days that match a search query. In the twitter query below, it searches
for recent tweets that contain the word "tariff".
```{r twitter-api}
# Retrieve bearer token from environment
bearer_token <- Sys.getenv("BEARER_TOKEN")
# Define token header and paste together (sprintf similar to paste0)
headers <- c(`Authorization` = sprintf('Bearer %s', bearer_token))
# Define parameters where max results is 500, querying for the string tariff, and then sorting results by relevance
params <- list(
    max_results = 100,
    query = "tariffs",
    sort_order = "relevancy"
    
)
# defining url_handle for the recent tweets url
url_handle <- "https://api.twitter.com/2/tweets/search/recent"
# saving the response object 
res <-
  httr::GET(url = url_handle,
            httr::add_headers(.headers = headers),
            query = params)

# Check response type
http_type(res)
# Check for errors
http_error(res)

# parsing the json
recent_search_body <-
  content(
    res,
    as = 'parsed',
    type = 'application/json',
    simplifyDataFrame = TRUE
  )

# df of tweet id, edit history of tweets, and text
recent_text_body <- recent_search_body$data
```

### Loading the Tweet Evalulation Model
The code below connects to the Roberta Base Sentiment model (documentation can be
seen [here](https://huggingface.co/cardiffnlp/twitter-roberta-base-sentiment?text=RT+%40Sojka4President%3A+You+know+who+doesn%E2%80%99t+have+to+pay+tariffs+because+they+manufacture+everything+in+the+United+States.+%5Cn%5CnHollywood+Product%E2%80%A6)).

By developing the function `get_sentiment`, an individual tweet sentiment can be
detected. 
```{r get-sentiment-funct}
# Retrieve bearer token from environment for HF Model 
hf_bearer_token <- Sys.getenv("HF_TOKEN")

# creating get_sentiment funct.
get_sentiment <- function(tweet){
# Get sentiment predictions from the Roberta Base Sentiment Model
  sent_res <- POST(
    url = "https://api-inference.huggingface.co/models/cardiffnlp/twitter-roberta-base-sentiment",
    add_headers("Authorization" = paste("Bearer", hf_bearer_token)),
    content_type("application/json"),
    body = toJSON(
      list(
        inputs = tweet
      ),
      auto_unbox = TRUE
    )
  )
  
sent_content <- fromJSON(rawToChar(sent_res$content))
}
```

The code below tests the sentiment for one tweet:
```{r testing-sentiment-funct}
# example tweet
tweet <- "You know who doesn't have to pay tariffs"
# checking that the funct above works
sentiment_result <- get_sentiment(tweet)
```

### Data Cleaning
To properly analyze tweets, the text has to be cleaned, and normalized so each
tweet can be properly compared. To clean the tweets, the function `clean_text`, 
is developed and applied to the 'recent_text_body' dataframe, creating a new
column in the dataframe titled 'clean_text'.
```{r clean-text-funct}
clean_text = function(rawtext)
{
  # convert to lower case
  rawtext = tolower(rawtext)
  # remove rt
  rawtext = gsub("rt", "", rawtext)
  # remove at
  rawtext = gsub("@\\w+", "", rawtext)
  # remove punctuation
  rawtext = gsub("[[:punct:]]", "", rawtext)
  # remove numbers
  rawtext = gsub("[[:digit:]]", "", rawtext)
  # remove links http
  rawtext = gsub("http\\w+", "", rawtext)
  # remove tabs
  rawtext = gsub("[ |\t]{2,}", "", rawtext)
  # remove blank spaces at the beginning
  rawtext = gsub("^ ", "", rawtext)
  # remove blank spaces at the end
  rawtext = gsub(" $", "", rawtext)
  # some other cleaning text
  rawtext = gsub('https://','',rawtext)
  rawtext = gsub('http://','',rawtext)
  rawtext = gsub('[^[:graph:]]', ' ',rawtext)
  rawtext = gsub('[[:punct:]]', '', rawtext)
  rawtext = gsub('[[:cntrl:]]', '', rawtext)
  rawtext = gsub('\\d+', '', rawtext)
  rawtext = str_replace_all(rawtext,"[^[:graph:]]", " ")
  return(rawtext)
}
# 
cleaned_recent_text <- recent_text_body |> 
  mutate(clean_text = clean_text(recent_text_body$text))
```

To get the sentiments for all tweets in the dataframe 'recent_text_body', another 
function is developed titled `get_sentiments_batch`, where each text goes through
a loop and the sentiment is recorded. A delay is incorporated into the function
to ensure the Roberta Base Sentiment API is not overwhelmed. 
```{r funct-all-cleaned-tweets}
# Create function with batching to apply get_sentiment() to all tweets
get_sentiments_batch <- function(tweets, ids, delay = 1) {
  results <- vector("list", length(tweets))  # create empty result list
  # iterate through each tweet
  for (i in seq_along(tweets)) {
    tweet <- tweets[i]
    id <- ids[i]
    # get the sentiment of the tweets 
    sentiment <- tryCatch({
      get_sentiment(tweet)
    }, error = function(e) {
      list(negative = NA, neutral = NA, positive = NA)
    })
    # record the sentiment results in the results list 
    results[[i]] <- c(id = id, sentiment)
    Sys.sleep(delay)
  }
  # bind the rows and save the results as a data frame
  return(do.call(rbind, lapply(results, as.data.frame)))
}

# incorporating 1 second delay; 0 = negative, 1= neutral, 2= positive
sentiment_results <- get_sentiments_batch(
                                          tweets = cleaned_recent_text$clean_text,
                                          ids = cleaned_recent_text$id,
                                          delay = 1
                                          )
```

Then, the dataframe is pivoted wider so that each observation is a tweet with 
the corresponding sentiment scores. 
```{r pivot-wider-sentiment-results}
# pivot wider so each observation is a tweet with a corresponding sentiment scores
sentiment_results_wider <- sentiment_results |> 
  pivot_wider(
    names_from = label, 
    values_from = score
    ) |> 
  left_join(recent_text_body |>  select(id, text), by = "id") |> 
  rename(
    "negative" = "LABEL_0",
    "neutral" = "LABEL_1",
    "positive" = "LABEL_2"
  )
```

To better understand the sentiment scores as a whole, the summarize function can 
be used to outline the mean, median, standard deviation, minimum and maximum 
values, and interquartile range of the negative, neutral, and positive scores.
```{r summary-stats}
# mean, median, sd, min, max, iqr of for each sentiment type
sentiment_roberta_base_summary <- sentiment_results_wider |> 
  summarize(
    mean_negative = mean(negative, na.rm = TRUE),
    median_negative = median(negative, na.rm = TRUE),
    sd_negative = sd(negative, na.rm = TRUE),
    min_negative = min(negative, na.rm = TRUE),
    max_negative = max(negative, na.rm = TRUE),
    iqr_negative = IQR(negative, na.rm = TRUE),
    mean_neutral = mean(neutral, na.rm = TRUE),
    median_neutral = median(neutral, na.rm = TRUE),
    sd_neutral = sd(neutral, na.rm = TRUE),
    min_neutral = min(neutral, na.rm = TRUE),
    max_neutral = max(neutral, na.rm = TRUE),
    iqr_neutral = IQR(neutral, na.rm = TRUE),
    mean_positive = mean(positive, na.rm = TRUE),
    median_positive = median(positive, na.rm = TRUE),
    sd_positive = sd(positive, na.rm = TRUE),
    min_positive = min(positive, na.rm = TRUE),
    max_positive = max(positive, na.rm = TRUE),
    iqr_positive = IQR(positive, na.rm = TRUE)
  )
```

The data is then pivoted wider so that each sentiment type and the proportion
of the sentiment type is recorded.
```{r sentiment-pivot-longer}
# pivoting longer to analyze all sentiment types
sentiment_long <- sentiment_results_wider %>%
  pivot_longer(cols = c(negative, neutral, positive),
               names_to = "sentiment_type",
               values_to = "proportion") |> 
  mutate(sentiment_type = case_when(
    sentiment_type == "negative" ~ "Negative",
    sentiment_type == "neutral"  ~ "Neutral",
    sentiment_type == "positive" ~ "Positive"
  )) |> 
  clean_names("title")
  
```

To better understand the distribution of the sentiment types for the tweets, a 
histogram can be used to visualize distribution by sentiment type.
```{r histogram}
ggplot(sentiment_long, aes(x = Proportion, fill = `Sentiment Type`)) +
  geom_histogram(binwidth = 0.05, color = "black", alpha = 0.7) +
  facet_wrap(~ `Sentiment Type`, ncol = 1, scales = "free_y") +
  theme_minimal() +
  labs(title = "Histogram of Sentiment Proportions",
       x = "Proportion", y = "Tweet Count") 
```

A density plot can also be used to better understand the distribution of the 
sentiment detected for the sample.
```{r}
ggplot(sentiment_long, aes(x = Proportion, color = `Sentiment Type`, fill = `Sentiment Type`)) +
  geom_density(alpha = 0.3) +
  theme_minimal() +
  labs(title = "Density Plot of Sentiment Proportions",
       x = "Proportion", y = "Density")
```
A boxplot may also be used to better understand the distribution, according to 
the plot below, neutral sentiment most frequently detected in the tweets.
```{r boxplots}
ggplot(sentiment_long, aes(x = `Sentiment Type`, y = Proportion, fill = `Sentiment Type`)) +
  geom_boxplot(alpha = 0.7) +
  theme_minimal() +
  labs(title = "Boxplot of Sentiment Proportions",
       x = "Sentiment Type", y = "Proportion")
```

### Sentiment Exploratory Analysis with `sentimentr` package
Another way to measure sentiment analysis is through the `sentimentr` package,
and by using the built in `sentiment_by` function. To compare the results from 
the `sentimentr` package and the Roberta Base Sentiment, a left_join() is used
to join the dataframes 'sentimentr_results' and 'roberta_net_sentiment' by the
tweet id. Then, a new column, titled 'roberta_net_sentiment' is created by 
taking the positive sentiment value minus the negative sentiment value.
```{r sentimentr}
# sentiment analysis with sentimentr package
sentimentr_results <- sentiment_by(cleaned_recent_text$clean_text, 
                                      by = cleaned_recent_text$id)

sentiment_compare <- sentimentr_results |> 
  left_join(sentiment_results_wider,
            by="id") |> 
  mutate(roberta_net_sentiment = positive - negative) |> 
  select(-sd)
```

To understand the relationship between the two senitment types, a scatterplot can 
be used to visualize the data:
```{r scatterplot-w-line-of-best-fit}
ggplot(sentiment_compare, aes(x = roberta_net_sentiment, y = ave_sentiment)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", color = "blue", se = FALSE) +
  scale_y_continuous(limits=c(-1,1))+
  theme_minimal() +
  labs(title = "Comparison of Roberta and sentimentr Scores",
       x = "Roberta Base Sentiment Result",
       y = "sentimentr Sentiment Result")
```

To better understand the correlation, the slope of the best fit line is 
calculated using the lm() function. The slope of a regression line indicates how 
the response variable changes on average for every unit change in the 
predictor variable. It also shows the relationsip of the variables. The positive
slope means that as the predictor variable increases, the response 
variable also increases. 
```{r calc-slope}
lm(data=sentiment_compare,
   formula = ave_sentiment~roberta_net_sentiment)
```

Another way to measure the relationship between the variables is the correlation.
```{r correlation}
cor(sentiment_compare$roberta_net_sentiment, sentiment_compare$ave_sentiment, use = "complete.obs")
```

## Conclusion
The results are slightly different. This means that further analysis and research
should be conducted to better understand how the models differ and which result
may be more accurate. 