---
title: "Using R and SQL to Analyze My Favorite Movies"
author: "Alyssa Gurkas"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: cerulean 
---

[R Pubs Link](https://rpubs.com/agurkas/data-607-assignment-2)
<br>
[Github Repository](https://github.com/amgurkas/data-acquisition) 

## Overview 
This R Markdown will explain how to connect to a MySQL Database, create a table,
and visualize the data.

```{r load-library}
library(RMySQL)
library(tidyverse)
library(ggplot2)
library(rstudioapi)
library(RColorBrewer)
```

## Data Collection
For this exercise, 12 survey recipients were contacted to rate six popular 
movies. The respondents were given instructions to rate the movies with five as
the highest rating, and to skip movies they had not seen.

In the code below, the RStudio API package is used to generate a pop-up message 
requesting the username and password for the MySQL database. However, you can
also use the .REnviron to save the respective username and password. 
```{r password-username-entry, include=TRUE}
username <- rstudioapi::askForPassword("Database username")
password <- rstudioapi::askForPassword("Database password")
```

```{r mysql-connection, include=FALSE}
mydb <- dbConnect(MySQL(), user = username, 
                  password = password, 
                  dbname = 'alyssa.gurkas84', 
                  host = "cuny607sql.mysql.database.azure.com")
```

The survey responses were collected in Google Forms, and were then exported to a 
.csv file. In the code below, the survey responses are read into R, and then used 
to create a MySQL table, with the connection established in the step above.
<br>
To better understand the dataset, the 'DESCRIBE' sql function can be used, 
additionally, the head() and glimpse() functions can be used to get a quick 
understanding of the dataset. 
```{r create-movie-database, include=TRUE}
movieRatingData <- read.csv(file="Movie_Rating_Responses.csv", 
                            header=TRUE, 
                            sep=",")

dbWriteTable(mydb, 
             value = movieRatingData, 
             row.names = FALSE, 
             name = "movie_ratings", 
             append = TRUE )

sql_table <- dbReadTable(mydb, "movie_ratings")

dbSendQuery(mydb, "DESCRIBE movie_ratings") |>  
  fetch()

head(sql_table)
glimpse(sql_table)

```

Then, to visualize the data, first it is helpful to know what the column names
are in the MySQL table, "movie_ratings". The column names are used to pivot the
dataframe longer, so we can create a bar graph depicting the received responses.

<br>

Using the 'ggplot2' package, bar charts for each of the movies are plotted, in 
one object. This can be accomplished by using the 'facet_wrap()' argument.
```{r plotting-data, echo=TRUE, include=TRUE}
colnames(sql_table)

movieRatingData_v2 <- sql_table |> 
  pivot_longer( cols = c("Shawshank.Redemption",
                         "The.Lion.King",
                         "Step.Brothers",
                         "Pulp.Fiction",
                         "Fight.Club"),
                names_to = "Movies", values_to = "Rating") |> 
  group_by(Movies,Rating) |> 
  summarize(Count=n()) |> 
  mutate(Movies=str_replace_all(Movies,"\\."," ")) 

ggplot(movieRatingData_v2)+
  geom_bar(aes(x=Rating,y=Count, fill=Movies),stat="identity")+
  scale_fill_brewer(palette = "Pastel1")+
  facet_wrap(~Movies)+
  theme(legend.position="none",
        strip.background = element_rect(
     color="grey", fill="white", size=.75, linetype="solid"))
  
```

## Note about missing data
Survey respondents were instructed to leave responses blank if they had not seen 
a movie. This meant that for some movies, there are missing data. For this
exercise, missing data was ignored. If more data about the respondents was 
collected, for example, demographic data, estimates for missing data may have 
been included through imputations.