---
title: "Simple Analysis of Bob Ross Paintings"
author: "Alyssa Gurkas"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: cerulean 
---

# R Pubs Link: https://rpubs.com/agurkas/data-607-assignment-1
# Github Repository: https://github.com/amgurkas/data-acquisition 

```{r load-library, echo=FALSE, include=FALSE, warnings=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
```

```{r load-data, echo=FALSE, include=FALSE, warnings=FALSE, message=FALSE}
data <- read_csv("data/elements-by-episode.csv")
```

## Overview

This report will conduct exploratory data analysis using data from 
FiveThirtyEight's 'elements-by-episode' dataset. This dataset tracks 67 
different elements found in Bob Ross paintings featured in the TV show, 
"The Joy of Painting". For more information about this dataset and 
a statistical analysis of the work of boss, please see [this article](https://fivethirtyeight.com/features/a-statistical-analysis-of-the-work-of-bob-ross/).

## Summary of the data
To better understand the dataset, we can run the following to return a vector 
showing how many rows and columns are within our dataset titled `data`:
```{r data-dim, echo=TRUE, include=TRUE}
dim(data)
```
To view summary statistics for the first six columns of the 
dataframe `data`, we run the following:

```{r summary-stats, echo=TRUE, include=TRUE}
summary(data[1:6])
```

For the columns that describe the paintings, the data are binary, meaning if
an element is present in the painting, the value will be recorded as 1. If the
element is not present, the value will be 0. Since there are 69 columns present,
it may be better to focus on a particular element, such as whether the paintings 
are framed. 

## Basic Exploratory Data Analysis
Currently, the data are at the episode level, meaning each row depicts the 
elements of a painting present in a given episode. For this report, we
are exploring which paintings are framed and will add a column that 
records which season the painting was featured in:
 
```{r basic-eda, echo=TRUE}
# creating a new column titled 'SEASON'
data_v2 <- data |> 
  mutate(SEASON = substr(EPISODE, 2,3)) 

# selecting columns that contain the term FRAME
data_v3 <- data_v2 |> 
  select(EPISODE,TITLE,SEASON, contains("FRAME"))
```

To calculate the percentage of total episodes that include a framed painting, 
and the percentage of total episodes that include an unframed painting we 
run the following:

```{r frame-stats, echo=TRUE, include=TRUE}
# calculating sum of framed paintings
sum(data_v3$FRAMED == 0)/sum(data_v3$FRAMED == 0|1)*100

# calculating sum of unframed paintings
sum(data_v3$FRAMED == 1)/sum(data_v3$FRAMED == 0|1)*100
```

```{r subset-frame, echo=FALSE, include=FALSE, warnings=FALSE, message=FALSE}
# data frame of framed paintings
data_v4 <- data_v3 |> 
  filter(FRAMED != 0) |> 
  select(-FRAMED)

frame_type <- data_v4 |> 
  select(contains("FRAME")) |> 
  colnames()
```

```{r painting-frame-analysis, echo=FALSE, include=FALSE, warnings=FALSE, message=FALSE}
data_v5 <- data_v4 |> 
  pivot_longer(
    cols= frame_type, names_to = "FRAME_TYPE", values_to = "INDICATOR"
  ) |> 
  group_by(SEASON,FRAME_TYPE) |> 
  summarise(COUNT = sum(INDICATOR))

data_v5 |> 
  filter(COUNT > 1)
```

```{r bar-chart-framed, echo=FALSE, include=TRUE, warnings=FALSE, message=FALSE}
bar_chart_df <- data_v3 |> 
  mutate(FRAMED = case_when(FRAMED == 0 ~"No Frame", 
                            FRAMED == 1 ~"Framed")) |> 
  select(EPISODE,SEASON,FRAMED,TITLE) |> 
  rename("Framed?" = "FRAMED")

ggplot(bar_chart_df, aes(x=`Framed?`, fill=`Framed?`))+
  geom_bar()+
  geom_text(stat='count', aes(label=after_stat(count)), vjust=-.25)+
  xlab("Painting Type")+
  ylab("Count")+
  ggtitle("Bob Ross Paintings by Frame")+
  theme_minimal()
```

## Findings 
Bob Ross did not typically frame his paintings. Frames were only featured in 
about 13% of the episodes. Additionally, Bob Ross did not feature a 
framed painting until Season 4, using a circle frame. When Bob Ross incorporated 
a frame for a featured painting, he most commonly used the oval frame. 
